Kids stove grades and usage summary

### 2020-03-16 pm_summary.Rmd
New summary plots

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 2 Feb 2020"
output: pdf_document
---

@@

pm_by_day <- pm_clean %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, day_of_week) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_day
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=5}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(0, 500) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

>>>>

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 16 March 2020"
output: powerpoint_presentation
---

@@

pm_by_day <- pm_clean %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, day_of_week) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
#pm_by_day
```

# Boxplots for PM by study area
```{r, fig.height=6, fig.width=8}
boxplots_kids_pm <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  ggplot(aes(area, pm)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=4, color="red",
                 position=position_nudge(x = -0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=mean, geom="text", 
                 size=7, position=position_nudge(x = -0.35, y = 0)) +
    stat_summary(fun.y=median, geom="point", shape=16, size=4, color="blue",
                 position=position_nudge(x = 0.1, y = 0)) +
    stat_summary(aes(label=round(..y.., digits = 1)), fun.y=median, geom="text", 
                 size=7, position=position_nudge(x = 0.35, y = 0)) +
    theme_minimal() +
    ylim(0, 500) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = expression(paste("PM"[2.5], " (", mu, g/m^3, ")"))) +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_pm
```

### 2020-03-16 stove_grade_usage_summary.Rmd
New summary plots

---
title: "Kids stove grades and usage summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 27 Jan 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

@@

# Summary of stove grades
```{r}
## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, stove_grade) %>% 
  #filter(!is.na(stove_grade)) %>% 
  group_by(area, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n())
kable(stove_grade_summary, align = "c")
```

## 18 homes with grades from 2 winters
## 4 of the homes were assigned different grades in each winter
### Was there an upgrade/downgrade to the stove? Check for comments/notes.
```{r}
multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
kable(multiple_grades)
```

# Summary of stove use
```{r}
# Summarize numeric vars
stove_use_summary <- stove_use_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
            "SD Cords" = sd(wood_cords, na.rm = TRUE),
            "Min Cords" = min(wood_cords, na.rm = TRUE), 
            "Median Cords" = median(wood_cords, na.rm = TRUE), 
            "Max Cords" = max(wood_cords, na.rm = TRUE))
kable(stove_use_summary, digits = 1, align = "c")

@@

# Stove use summary plots
```{r}
stove_use_plot_function <- function(data, variable) {

stove_use_plot <- data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

@@

stove_use_plot_function(stove_use_clean, "burn_level")
stove_use_plot_function(stove_use_clean, "wood_type")
stove_use_plot_function(stove_use_clean, "wood_collect")
stove_use_plot_function(stove_use_clean, "wood_collect_method")
```

#### Abstract/manuscript tables ####

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

@@

# Summarize home data
```{r}
# Summarize numeric vars
kids_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
            "SD" = sd(wood_cords, na.rm = TRUE),
            "min" = min(wood_cords, na.rm = TRUE), 
            "med" = median(wood_cords, na.rm = TRUE), 
            "max" = max(wood_cords, na.rm = TRUE))
kable(kids_use, digits = 1, align = "c")

>>>>

---
title: "Kids stove grades and usage summary"
author: "Ethan Walker"
date: "Started 15 Jan 2020, Updated 16 March 2020"
output: pdf_document 
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
stove_grades_clean <- read_rds("Output/stove_grades_clean.rds")
stove_use_clean <- read_rds("Output/stove_use_clean.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds") 
demographics_clean <- read_rds("Output/demographics_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")
```

@@

```{r}
## Look further into NA values? ##
stove_grade_summary <- stove_grades_clean %>% 
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, home, stove_grade, winter_id) %>% 
  #filter(!is.na(stove_grade)) %>% 
  group_by(area, winter_id, stove_grade) %>% 
  distinct(home) %>% 
  summarize(n()) %>% 
  filter(!is.na(stove_grade))
kable(stove_grade_summary, align = "c")

stove_grades_total <- stove_grades_clean %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, stove_grade) %>% 
  arrange(area, home) 
```

```{r}
# Generate random sample of stoves to re-grade and validate grading methods

# 10% of homes from each area, earlier winters
# send new ID linked to new ID to Kathrene

random_sample_wmt <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  mutate(winter_id = as.factor(winter_id)) %>% 
  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  filter(cutoff_date == 1) %>% 
  filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
  arrange(area, home) %>% 
  group_by(area) %>% 
  sample_n(17)
#write_csv(random_sample_wmt, "Output/random_sample_wmt.csv")

random_sample_nn <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  mutate(winter_id = as.factor(winter_id)) %>% 
  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  filter(cutoff_date == 1) %>% 
  filter(area == "NN") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
  arrange(area, home) %>% 
  group_by(area) %>% 
  sample_n(8)
#write_csv(random_sample_nn, "Output/random_sample_nn.csv")

random_sample_ak <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  mutate(winter_id = as.factor(winter_id)) %>% 
  left_join(demographics_clean, c("area", "home", "winter_id")) %>% 
  mutate(cutoff_date = if_else(intervention_date < "2018-06-06", 1, 0)) %>% 
  filter(!is.na(stove_grade)) %>% 
  filter(!is.na(home)) %>% 
  filter(cutoff_date == 1) %>% 
  filter(area == "AK") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, stove_grade, intervention_date) %>% 
  arrange(area, home) %>% 
  group_by(area) %>% 
  sample_n(5)
#write_csv(random_sample_ak, "Output/random_sample_ak.csv")
```

```{r, include=FALSE}
## 18 homes with grades from 2 winters
## 4 of the homes were assigned different grades in each winter
### Was there an upgrade/downgrade to the stove? Check for comments/notes.

multiple_grades <- stove_grades_clean %>% 
  select(area, home, stove_grade, winter_id) %>% 
  filter(winter_id == 1 | winter_id == 2) %>% 
  filter(!is.na(stove_grade)) %>%
  group_by(home) %>% 
  filter(n() > 1) %>% 
  arrange(home, winter_id)
kable(multiple_grades)
```

```{r, include=FALSE}
# Summarize numeric vars
stove_use_summary <- stove_use_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area) %>% 
  summarize("N" = n(),
            "Mean Cords" = mean(wood_cords, na.rm = TRUE),
            "SD Cords" = sd(wood_cords, na.rm = TRUE),
            "Min Cords" = min(wood_cords, na.rm = TRUE), 
            "Median Cords" = median(wood_cords, na.rm = TRUE), 
            "Max Cords" = max(wood_cords, na.rm = TRUE))
kable(stove_use_summary, digits = 1, align = "c")

@@

```{r}
stove_grade_plot <- stove_grades_clean %>% 
  filter(winter_id == 1) %>% 
  select(area, stove_grade) %>% 
  filter(!is.na(stove_grade)) %>% 
  ggplot() +
    geom_bar(aes(stove_grade, fill = area)) +
    theme_minimal() +
    labs(title = "Stove grades") +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_grade_plot
```

```{r}
stove_use_data <- read_rds("Output/stove_use_clean.rds") %>% 
  mutate(wood_cords = as.factor(wood_cords))

stove_use_plot_function <- function(data, variable) {

stove_use_plot <- data %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  select(area, variable) %>% 
  rename(plot_var = variable) %>% 
  #na.exclude() %>% 
  ggplot() +
    geom_bar(aes(plot_var, fill = area)) +
    theme_minimal() +
    labs(title = variable) +
    theme(axis.text.x = element_text(size = 16, colour = "black", 
                                     angle = 40, vjust = 0.7, hjust = 0.7),
          axis.text.y = element_text(size = 16, colour = "black"),
          title = element_text(size = 18, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    scale_fill_manual(values = jv_palette)
stove_use_plot

@@

stove_use_plot_function(stove_use_data, "burn_level")
stove_use_plot_function(stove_use_data, "wood_type")
stove_use_plot_function(stove_use_data, "wood_collect")
stove_use_plot_function(stove_use_data, "wood_collect_method")
stove_use_plot_function(stove_use_data, "wood_cords")
```

```{r}
#### Abstract/manuscript tables ####

# Filter for homes that have PM data

# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

@@

```{r, include=FALSE}
# Summarize numeric vars
kids_use <- stove_use_filtered %>%  
  #distinct(home, .keep_all = TRUE) %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(wood_cords, na.rm = TRUE), 
            "SD" = sd(wood_cords, na.rm = TRUE),
            "min" = min(wood_cords, na.rm = TRUE), 
            "med" = median(wood_cords, na.rm = TRUE), 
            "max" = max(wood_cords, na.rm = TRUE))
kable(kids_use, digits = 1, align = "c")
